### 2016-10-22 paper.tex
writing section 3

\documentclass[10pt]{sigplanconf}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{listings}
\usepackage[noline]{algorithm2e}
\usepackage{comment}
\usepackage{color}
\usepackage{times}
\usepackage{multirow}
\usepackage{balance}
\usepackage{framed}
\usepackage{alltt}
\usepackage{textcomp}
\usepackage[dvipsnames]{xcolor}

\setcounter{tocdepth}{3}

% configures style of source code listings
\lstset{
	basicstyle=\small\sffamily
	%basicstyle=\scriptsize\ttfamily
	%numbers=left
}

\lstdefinelanguage{java}{
	morekeywords={abstract,continue,for,new,switch,assert,default,goto,package,synchronized,boolean,stream,event,filter,when,do,if,private,this,break,double,implements,protected,throw,byte,else,import,public,throws,case,enum,instanceof,return,transient,catch,extends,int,short,try,char,final,interface,static,void,class,finally,long,strictfp,volatile,const,float,native,super,while},
	sensitive=false,
	morestring=[b]",
	morestring=[b]',
	morecomment=[l]{//},
	morecomment=[s]{/*}{*/)},
	escapeinside=??,
	moredelim=[is][\textit]{__}{__},
	moredelim=[is][\textbf]{_*}{*_}
}
\lstnewenvironment{displayjava}
	{\lstset{language=java,basicstyle=\small\sffamily,tabsize=2,columns=fullflexible,captionpos=b,xleftmargin=1em,xrightmargin=1em}}{}
\lstnewenvironment{java}
	{\lstset{language=java,basicstyle=\small\sffamily,tabsize=2,columns=fullflexible,captionpos=b}}{}

\newcommand{\code}[1]{{\sf \small #1}\xspace}

\newcommand{\mi}[1]{\mathit{#1}}
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\mt}[1]{\mathtt{#1}}

\newcommand{\todo}[1]{{\bfseries TODO #1}}

% for writing comments
\newcommand{\PP}[1]{\textcolor{blue}{\bfseries PP: #1}} % Pavel Parizek
\newcommand{\PV}[1]{\textcolor{red}{\bfseries PV: #1}} % Premek Vysoky
\newcommand{\JB}[1]{\textcolor{orange}{\bfseries JB: #1}} % JetBrains

% coloring and font style commands
  % ANTLR
\newcommand{\antlrap}{\textquotesingle}
\newcommand{\antlrparserrule}[1]{\textcolor{Blue}{#1}}
\newcommand{\antlrlexerrule}[1]{\textcolor{RedOrange}{#1}}
\newcommand{\antlrliteralnoap}[1]{\textcolor{Maroon}{#1}}
\newcommand{\antlrliteral}[1]{\antlrap\antlrliteralnoap{#1}\antlrap}
\newcommand{\antlrregex}[1]{\textcolor{PineGreen}{#1}}
  % MPS in general
\newcommand{\mpsconcept}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\mpsinterface}[1]{\textcolor{Plum}{#1}}
  % MPS structure
\newcommand{\mpsstkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpsstalias}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpsstproperty}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpsstplaceholder}[1]{\textcolor{Gray}{\textbf{#1}}}
\newcommand{\mpsstcardinality}[1]{\textcolor{RoyalBlue}{#1}}
  % MPS editor
\newcommand{\mpsedkeyword}[1]{\textcolor{Black}{\textbf{#1}}}
\newcommand{\mpsedtarget}[1]{\textcolor{Fuchsia}{#1}}
\newcommand{\mpsedannotation}[1]{\textcolor{Gray}{\textbf{#1}}}
  % MPS textgen
\newcommand{\mpstgkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpstgtarget}[1]{\textcolor{Fuchsia}{#1}}
\newcommand{\mpstgparam}[1]{\textcolor{Violet}{#1}}
\newcommand{\mpstgaction}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpstgliteral}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpstgnodeprop}[1]{\textcolor{darkgray}{#1}}

\begin{document}

% TODO
%\conferenceinfo{PLDI'17,} {June 19--26, 2017, Barcelona, Spain.}
%\CopyrightYear{2017}
%\copyrightdata{TODO}

\title{INGRID: Creating Languages in MPS from ANTLR Grammars}

\authorinfo{P\v{r}emysl Vysok\'y \and Pavel Par\'izek}
	{Charles University, Faculty of Mathematics and Physics, Prague, Czech Republic}
	{premek.vysoky@gmail.com, parizek@d3s.mff.cuni.cz}

% TODO pridat nekoho dalsiho z JetBrains (Alex Shatalin?)
% TODO upravit affiliation pro JetBrains -> doplnit "s.r.o" nebo radsi "Inc."
% TODO muzeme taky napsat jejich emaily pokud budou lidi z JetBrains souhlasit
\authorinfo{V\'aclav Pech}
	{JetBrains}

\maketitle

\begin{abstract}
% TODO
\todo{write later}

\todo{Copy from master thesis.}
JetBrains MPS is a language workbench focusing on domain-specific languages.
Unlike many other language workbenches and IDEs, it uses a projectional editor for code.
The developer directly manipulates the program in its tree form (AST) and not by editing a text source code.
This brings many advantages, but on the other hand requires time-consuming and complicated MPS language definition.
The thesis elaborates on the possibility of automating the process of creating MPS language definition from its grammar description.
It introduces the MPS editor, evaluates approaches of related projects and describes author's efforts to implement an MPS plugin that allows this import.
The chosen approach and the selection of tools used for implementation are justified in the thesis.
We point out important problems that any similar project might deal with and we introduce some possible solutions.
Furthermore, the thesis contains examples of imported languages, showing the potency of the chosen approach.
The thesis also aims to lay groundwork for future extensions and suggest possible improvements.
\end{abstract}

% TODO possible keywords: JetBrains MPS, programming language, grammar, projectional editor

\input{chapters/introduction}
\input{chapters/background}
\input{chapters/language_definition}
\input{chapters/evaluation}
\input{chapters/conclusion}

\acks
% TODO
This research was supported by JetBrains. \todo{Pridat jeste nejaky grant a kompletne preformulovat.}

\bibliographystyle{abbrvnat}

\begin{thebibliography}{}

\bibitem{ref:BISONFLEX} J. Levine. Flex \& Bison: Text Processing Tools. O'Reilly Media, 2009.

\bibitem{ref:ANTLRBOOK} T. Parr. The Definitive ANTLR 4 Reference. Pragmatic Bookshelf, 2013.

\bibitem{ref:MBEDDR} M. Voelter, D. Ratiu, B. Schaetz, and B. Kolb. mbeddr: an Extensible C-based Programming Language and IDE for Embedded Systems. SPLASH 2012, ACM.

\bibitem{ref:ANTLR} ANTLR parser generator, \url{http://www.antlr.org}

\bibitem{ref:ANTLR2MPS} F. Campagne. The ANTRL{\_}MPS project, \url{https://github.com/CampagneLaboratory/ANTLR_MPS}

\bibitem{ref:MPSMETABNF} DSLFoundry. The mps-metabnf project, \url{https://github.com/DSLFoundry/mps-metabnf}

\bibitem{ref:MPS} JetBrains MPS, \url{https://www.jetbrains.com/mps/}

\bibitem{ref:PE4MPS} M. Lombardo. The PE4MPS project, \url{https://github.com/mar9000/pe4mps}

\bibitem{ref:PE} M. Lombardo. PE grammars, \url{https://github.com/mar9000/pe}

\bibitem{ref:BASELANG} MPS BaseLanguage, \url{https://confluence.jetbrains.com/display/MPSD34/Base+Language}

% TODO opravit formatovani nasledujicich polozek a vhodne zaradit

\bibitem{ref:ECMAScript4MPS} ECMAScript4MPS project, \url{https://github.com/mar9000/ecmascript4mps}

\bibitem{ref:javascript} JavaScript ANTLRv4 grammar, \url{https://github.com/antlr/grammars-v4/blob/master/ecmascript/ECMAScript.g4}

\end{thebibliography}

\end{document}