### 2016-11-09 chapters/language_definition.tex
small changes

\section{Creating Language Definition in MPS}
\label{sect:LANGDEF}

The proposed \textsc{Ingrid} method accepts grammars in the ANTLR v4 notation~\cite{ref:ANTLRBOOK} as input.
All widely used programming languages have their syntax defined using this notation~\footnote{https://github.com/antlr/grammars-v4}.
Unlike some of the related projects, we did not extend the notation with any custom features, and we also did not create an MPS language for the ANTLR notation.

The process of MPS language construction, as performed by \textsc{Ingrid}, consists of four phases --- the first is parsing of the input grammar, which is followed by definition of the essential aspects of the language (Structure, Editor, and TextGen).
Each of the phases 2-4 is exclusively and fully responsible for one aspect.

\textsc{Ingrid} is currently able to create (i) a full Structure aspect for each element (type of an AST node) of the given language, (ii) a very basic Editor, and (iii) a basic TextGen aspect.
Therefore, the resulting MPS language typically still has to be adjusted manually to improve its usability, and the human users must also define the remaining aspects not yet supported by \textsc{Ingrid}.

Our approach differs from the related projects (Section~\ref{sect:RELATED}) especially in the level of automation. It also has much better support for the definition of Editor and TextGen.

\subsection{Running Example}

We will describe the \textsc{Ingrid} method using the example of a simplified XML language that is defined in Figure~\ref{fig:SIMPLEXML}.
The \emph{SimpleXML} language is small but complex enough to be used for illustration of the main challenges and behavior of the proposed algorithms.

\begin{figure*}[ht]
\centering
\begin{framed}
\begin{alltt}
\textbf{grammar} \textit{SimpleXML};
\antlrparserrule{document}  : \antlrparserrule{prolog}? \antlrparserrule{comment}? \antlrparserrule{element} ;
\antlrparserrule{prolog}    : \antlrliteral{<?xml } \antlrparserrule{attrib}* \antlrliteral{?>} ;
\antlrparserrule{comment}   : \antlrliteral{<!--} \antlrlexerrule{TEXT} \antlrliteral{-->} ;
\antlrparserrule{element}   : \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{>} \antlrparserrule{content}* \antlrliteral{</} \antlrlexerrule{Name} \antlrliteral{>} | \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{/>} ;
\antlrparserrule{attrib}    : \antlrlexerrule{Name} \antlrliteral{="} \antlrlexerrule{TEXT} \antlrliteral{"} ;
\antlrparserrule{content}   : \antlrlexerrule{TEXT} | \antlrparserrule{element} | \antlrparserrule{comment} | \antlrlexerrule{CDATA} ;
\antlrlexerrule{Name}      : \antlrlexerrule{NameStartChar} \antlrlexerrule{NameChar}* ;
\antlrlexerrule{DIGIT}     : \antlrregex{[0-9]} ;
\antlrlexerrule{NameChar}  : \antlrlexerrule{NameStartChar} | \antlrliteral{-} | \antlrliteral{\_} | \antlrliteral{.} | \antlrlexerrule{DIGIT} ;
\antlrlexerrule{NameStartChar} : \antlrregex{[:a-zA-Z]} ;
\antlrlexerrule{TEXT}      : \antlrregex{~[<"]*} ;
\antlrlexerrule{CDATA}     : \antlrliteral{<![CDATA[} \antlrregex{.*?} \antlrliteral{]]>} ;
\end{alltt}
\end{framed}
\caption{Grammar of the SimpleXML language in the ANTLR v4 notation}
\label{fig:SIMPLEXML}
\end{figure*}

The grammar of SimpleXML contains elements of the kinds that are listed below.
Each color in Figure~\ref{fig:SIMPLEXML} corresponds to a different kind in a way that is indicated by the list item headers.
\begin{itemize}
	\item \textbf{ANTLR v4 keywords} are required by the notation.
	\item \antlrparserrule{\textbf{Parser rules}} describe the structure of a language.
		Alternatives on the right side of a rule are separated by the pipe character ($|$).
	\item \antlrlexerrule{\textbf{Lexer rules}} describe terminal symbols that the parser matches against the input string.
		A terminal symbol can be encoded as a string value or using a regular expression.
	\item \antlrliteralnoap{\textbf{String literals}}, always written inside of a pair of single quote marks, also represent terminal symbols, but by exact match to a string constant.
	\item \antlrregex{\textbf{Regular expressions}} describe string tokens to be matched against regular expressions with a special ANTLR v4 regex notation\footnote{https://github.com/antlr/antlr4/blob/master/doc/lexer-rules.md}.
\end{itemize}
Elements on the right side of a rule can be annotated with standard EBNF operators (\code{?}, \code{+}, \code{*}) that specify the allowed number of occurrences.

\subsection{Phase 1: Parsing Input Grammar}

The first phase in the process of MPS language construction is parsing of the input ANTLR v4 grammar.
It is done using an ANTLR parser that was automatically generated from the grammar of the ANTLR v4 notation itself.
Nevertheless, the full AST, which comes out of the parser, is quite complex and contains information not relevant for the \textsc{Ingrid} algorithm.
In order to get a simple representation that is easy to process by the later phases and keeps only information necessary for the construction of MPS languages, several steps of post-processing of the AST are performed in this phase.
An output is a custom AST-like structure representing the grammar, especially the hierarchy of parser rules.

The representation of lexer rules (tokens) in the AST has to be simplified too.
Note that, in the ANTLR notation, lexer rules can be built from alternatives just like the parser rules --- see, for example, the lexer rule \antlrlexerrule{Name} from our SimpleXML language (Figure~\ref{fig:SIMPLEXML}).
For each lexer rule, the parser produces a tree that captures the structure of the rule.
However, the lexer rules are, in fact, just regular expressions used to recognize tokens in the input string.

Every tree that represents a lexer rule is flattened into the equivalent regular expression by the recursive algorithm specified in Figure~\ref{fig:ALGFLATTEN}.
A distinct sequence is created from the elements of each alternative, and then all those sequences are joined by the alternation operator \code{$|$} to form a regular expression.

\begin{figure}[ht]
\centering
\begin{framed}
\begin{alltt}
Flatten(\textit{R}):
  \textit{T} = empty list
  \textbf{for each} alternative \textit{A} \textbf{in} rule \textit{R}:
    \textit{R} = \antlrap\antlrap
    \textbf{for each} element \textit{E} \textbf{of} \textit{A}:
      \textbf{if} \textit{E} is not yet flattened \textbf{then}:
        Flatten(\textit{E})
      \textit{R}.append(\textit{E})
      \textit{R}.append(\textit{E}.operator)
    \textit{T}.add(\textit{R})
  build string S from elements of \textit{T}:
    \textit{S} = \(t\sb{1}\) | \(t\sb{2}\) | \(\ldots\) | \(t\sb{n}\)
  \textbf{return} \textit{S}
\end{alltt}
\end{framed}
\caption{Flattening algorithm}
\label{fig:ALGFLATTEN}
\end{figure}

The output of \texttt{Flatten(Name)}, i.e. application of the algorithm to the \antlrlexerrule{Name} rule, is this regular expression:

\begin{center}
  \texttt{[:a-zA-Z]([:a-zA-Z]|\textbackslash-|{\_}|\textbackslash.|[0-9])*}
\end{center}
It defines the syntactically valid identifiers of elements and attributes in SimpleXML.
Each identifier must begin with a letter, followed by any combination of letters, digits, underscore, dash, or a dot.

\subsection{Phase 2: Structure}

In the next phase, the complete structure of the MPS language is automatically generated from the AST that represents syntax of the input language.
Elements of the Structure aspect are derived from the AST nodes, and then linked appropriately.
Therefore, structure of the MPS language is usually similar to the original ANTLR grammar.

When designing the process for translating AST nodes (i.e., grammar rules) into the MPS language structure, we faced several challenges.
The main challenge related to the structure, as we already mentioned in the introduction, is that a grammar typically contains rules that do not directly correspond to programming language constructs.
Such rules exist at the intermediate layers of a syntax hierarchy.
Their main purpose is to enable easier understanding and maintenance of the grammar by humans.
However, presence of the intermediate layers significantly complicates usage of the given language in MPS, and the layers are actually not necessary for construction of an MPS language.
We show examples illustrating this problem, which we call a \emph{layer problem}, later in the subsection where we describe our solution.

As a part of the \textsc{Ingrid} method, we have designed an approach to eliminate the unnecessary layers during construction of the Structure aspect --- we call it the \emph{shortcut} approach (Section~\ref{sect:SHORTCUT}).
However, before focusing on the intermediate layers and the shortcut approach, we describe the basic principles of translation from the AST into the language structure.

In MPS, each concept of the language (i.e., every AST node) is represented by an object that may have a parent, some children, and properties of any data type.
Additionally, the object may implement any number of interfaces, and it may also contain references to objects representing other AST nodes.
The parent-child relationships between objects that make the Structure aspect are derived from rules of the grammar (i.e., from the structure of the AST).
For each rule, the object corresponding to the left-hand side is in the parent-child relationships with sets of objects that represent language elements referenced by the right-hand side.
If a rule has multiple alternatives, then a distinct object (MPS concept) has to be created in the structure for each alternative.
The name of a concept (object) in MPS is composed from (1) the name of the AST node, which is equivalent to the string encoding of the left-hand side of the corresponding rule, and (2) the number indicating the position of the respective alternative on the right-hand side of the rule.

\begin{figure}[ht]
\centering
\begin{alltt}
\small
\mpsstkeyword{concept} Element\_1 \mpsstkeyword{extends} BaseConcept
        \mpsstkeyword{implements} IContent, IElement

  \mpsstkeyword{instance can be root:} false
  \mpsstkeyword{alias:} \mpsstalias{< > </ >}
  \mpsstkeyword{short description:} \mpsstalias{Element}

  \mpsstkeyword{properties:}
  \mpsstproperty{Name\_1} : Name
  \mpsstproperty{Name\_2} : Name

  \mpsstkeyword{children:}
  \mpsstproperty{Attribute\_1} : Attribute[\mpsstcardinality{0..n}]
  \mpsstproperty{Content\_2}   : IContent[\mpsstcardinality{0..n}]
\end{alltt}
\caption{Structure aspect of the \code{Element{\_}1} concept}
\label{fig:ELEMENTSTRUCT}
\end{figure}

Consider the parser rule \antlrparserrule{element} from the grammar in Figure~\ref{fig:SIMPLEXML}.
Its first alternative represents the full XML element with content.
Figure~\ref{fig:ELEMENTSTRUCT} shows the Structure aspect for the MPS concept (object) named \code{Element{\_}1} that corresponds to the alternative.
The object contains two properties, one for each reference to the lexer rule \antlrlexerrule{Name}.
Values of these properties will be restricted using the regular expression that corresponds to the lexer rule \antlrlexerrule{Name}.
String literals, such as the opening and closing brackets (\antlrliteral{\textless}, \antlrliteral{/\textgreater}) in XML, are omitted because they will be defined only in the Editor aspect for this concept.

References to other parser rules are captured by pointers to child objects.
The types of child objects, such as \code{IContent} in the \code{Element{\_}1} concept, are determined as follows.
Consider the \antlrparserrule{content} rule from the SimpleXML language.
We show just the relevant fragment of the grammar again in Figure~\ref{fig:CONTENTRULE}.
An object corresponding to any one of the four alternatives could be the actual value anywhere the \antlrparserrule{content} rule is referenced.

\begin{figure}[ht]
\centering
\begin{alltt}
\small
\antlrparserrule{content} : \antlrlexerrule{TEXT} | \antlrparserrule{element} | \antlrparserrule{comment} | \antlrlexerrule{CDATA} ;
\end{alltt}
\caption{Parser rule \antlrparserrule{content}}
\label{fig:CONTENTRULE}
\end{figure}

Our solution is to use interface concepts.
For each rule with more than one alternative on the right side, first an interface concept is defined in the MPS language structure, and then one object that implements the given interface is created for each alternative.
The resulting fragment of the language structure looks like the one in Figure~\ref{fig:ICONTENTITF}.
It contains one interface \mpsinterface{IContent} that is implemented by four object concepts \mpsconcept{Content{\_}1}, $\ldots$, \mpsconcept{Content{\_}4}.

\begin{figure}[ht]
\centering
\begin{alltt}
\small
\mpsinterface{IContent} : \mpsconcept{Content{\_}1} | \mpsconcept{Content{\_}2} |
           \mpsconcept{Content{\_}3} | \mpsconcept{Content{\_}4}
\end{alltt}
\caption{MPS interface \code{IContent} and the types of implementing objects}
\label{fig:ICONTENTITF}
\end{figure}

Now we can illustrate the layer problem on the behavior of auto-completion in MPS.
Suppose that a user is creating a new document in the SimpleXML language, just inserted a fresh node of the type \mpsconcept{Element{\_1}} (see above), and would like to insert another nested XML element inside.
The auto-completion mechanism of MPS offers four options that are displayed in the left part of Figure~\ref{fig:LAYERPROBLEM}.
Each option represents one of the MPS concepts that implement the interface \mpsinterface{IContent}, and therefore also one alternative of the \antlrparserrule{content} rule.

\begin{figure*}[ht]
	\centering
	\includegraphics[scale=0.5]{./images/layer_problem.png}
	\caption{Layer problem in auto-completion}
	\label{fig:LAYERPROBLEM}
\end{figure*}

However, in order to correctly insert another nested element, a user has to perform two steps:
\begin{enumerate}
	\item Insert an object (node) of the type \mpsconcept{Content{\_}2} inside the \mpsconcept{Element{\_}1} node. The \mpsconcept{Content{\_}2} object has only a single child node of the interface type \mpsinterface{IElement}.
	\item Then, the user must trigger the auto-complete again and insert either an \mpsconcept{Element{\_}1} node or an \mpsconcept{Element{\_}2} node into the \mpsconcept{Content{\_}2} node. See the right part of Figure~\ref{fig:LAYERPROBLEM}.
\end{enumerate}
The main difficulty here is that, in the first step, the user has to (i) either correctly guess what option offered by the auto-complete menu to select, or (2) to remember the order of alternatives in the grammar rule.

Similarly, if the user would like to replace a nested \mpsconcept{Element{\_}1} node with, for example, an XML comment (represented by the \mpsconcept{Comment} node), then both intermediary layers have to be deleted before she gets back to the original selection among the concepts \mpsconcept{Content{\_}1}, $\ldots$, \mpsconcept{Content{\_}4}.

Intermediary layers have no visual appearance, and therefore it is very difficult for users to see what is actually happening and they may get confused very easily.
The layer problem is addressed by the shortcut approach, which we describe in the next subsection.

\subsubsection{The Shortcut Approach}
\label{sect:SHORTCUT}

The key idea of this approach is to skip all the intermediary layers (nodes) in the syntax tree and consider just the leaf nodes, e.g., to be directly offered to the user through the auto-completion menu.
Specifically, the \antlrparserrule{content} rule from the SimpleXML grammar, which is given in Figure~\ref{fig:CONTENTHIERARCHY} together with rules that determine the relevant fragment of the syntax hierarchy, expands ultimately into the leaf nodes that are highlighted using the bold font in Figure~\ref{fig:CONTENTEXPAND}.

\begin{figure*}[ht]
\centering
\begin{framed}
\begin{alltt}
  \antlrparserrule{content} : \antlrlexerrule{TEXT} | \antlrparserrule{element} | \antlrparserrule{comment} | \antlrlexerrule{CDATA} ;
  \antlrparserrule{element} : \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{>} \antlrparserrule{content}* \antlrliteral{</} \antlrlexerrule{Name} \antlrliteral{>} | \antlrliteral{<} \antlrlexerrule{Name} \antlrparserrule{attrib}* \antlrliteral{/>} ;
  \antlrparserrule{comment} : \antlrliteral{<!--} \antlrlexerrule{TEXT} \antlrliteral{-->} ;
\end{alltt}
\end{framed}
\caption{The parser rule \antlrparserrule{content} with other rules that determine the relevant fragment of the syntax hierarchy}
\label{fig:CONTENTHIERARCHY}
\end{figure*}

\begin{figure}[ht]
\begin{framed}
\textbf{Content{\_}1} (TEXT) \\
\ \ \ Content{\_}2 $\rightarrow$ \textbf{Element{\_}1} \\
\ \ \ Content{\_}2 $\rightarrow$ \textbf{Element{\_}2} \\
\ \ \ Content{\_}3 $\rightarrow$ \textbf{Comment} \\
\ \ \ \textbf{Content{\_}4} (CDATA)
\end{framed}
\caption{Leaf nodes of the parser tree fragment that has the \antlrparserrule{content} rule as its root}
\label{fig:CONTENTEXPAND}
\end{figure}

For the input that consists of a particular AST node $N$ and the grammar rule $R$ that expands $N$, the procedure implementing the shortcut approach systematically traverses the parser tree (AST) built in the earlier phases in order to identify each AST node that may appear at the end of some derivation chain starting by the rule $R$ from the node $N$.
Such nodes cannot expand further based on any rule in the grammar, and for that reason we call them \emph{end nodes}.

\begin{figure*}[ht]
\begin{framed}
\begin{alltt}
 1 FindAllPathsToEndNodes(\textit{R}):
 2   \textit{CurPath} = empty list of rules and nodes
 3   \textbf{return} FindPaths(\textit{R}, \textit{CurPath})
 4
 5 FindPaths(\textit{R}, \textit{CurPath}):
 6  \textit{Paths} = empty list
 7  \textbf{for each} alternative \textit{A} \textbf{in} rule \textit{R}:
 8    \textit{NewCurPath} = Clone(\textit{CurPath})
 9    \textbf{if} \textit{A} contains only a single element \textit{E}:
10      \textit{NewCurPath}.Add(\(N\sb{E}\)) \textbf{where} \(N\sb{E}\) is the node representing \textit{E}
11      \textit{P} = FindPaths(\(R\sb{E}\), \textit{NewCurPath}) \textbf{where} \(R\sb{E}\) is the rule that expands \textit{E}
12      \textit{Paths} = Merge(\textit{Paths}, \textit{P})
13    \textbf{else}:
14      \textit{NewCurPath}.Add(\textit{R})
15      \textit{NewCurPath}.Add(\(N\sb{A}\)) \textbf{where} \(N\sb{A}\) is the node representing \textit{A}
16      \textit{Paths}.Add(\textit{NewCurPath})
17  \textbf{return} \textit{Paths}
\end{alltt}
\end{framed}
\caption{Algorithm to find all paths to end nodes for a parser rule}
\label{fig:SHORTCUTALG}
\end{figure*}

Figure~\ref{fig:SHORTCUTALG} shows the algorithm that finds all paths to some end node from a given parser rule $R$.
The algorithm is based on recursive traversal of the parser tree.
At each level of recursion, it gathers all paths that lead from the current parser rule $R$ to some end node through its alternatives (line~7).
Two cases may occur:
\begin{itemize}
	\item When a particular alternative $A$ of the rule $R$ contains only a single element $E$, and the element is a reference to another parser rule (line 9), the alternative $A$ is an intermediary layer that can be transparently hidden from the user of the MPS language.
		  A run of the algorithm continues, at line 11, by recursively processing alternatives of the rule corresponding to $E$, which lies at the next level of the parser tree.
	\item Otherwise, an end node of a derivation chain was found and the recursion stops (lines 13-16).
\end{itemize}
By appending the node corresponding to the current alternative (lines 10 and 15) and the rule that leads to the alternative (line 14), the algorithm creates a full path that contains the target end node as the last element of the chain.

We use the name \emph{shortcut approach} for this algorithm, because the paths collected by the algorithm provide shortcuts from the given rule to end nodes, by the virtue of hiding all intermediary layers.
For example, the result of this algorithm for the \antlrparserrule{content} rule (Figure~\ref{fig:CONTENTHIERARCHY}) is the list of five items already shown in Figure~\ref{fig:CONTENTEXPAND}.

The primary use case for shortcuts is to generate options for the auto-completion menus, such that only the end nodes are offered.
Shortcuts have to be considered not only when nodes are inserted, but also when they are deleted.
In each case, the whole chain including possibly multiple intermediary AST nodes (up to the end node) must be added, respectively deleted.
When the user wants to delete some end node from the AST of a program or document written in the MPS language, the effect of an insertion must be fully reversed.

\subsection{Phase 3: Editor}
\label{sect:EDITORDEF}

Having the complete structure of the new MPS language, the next phase is to define the visual representation of all concepts (AST nodes) in the projectional editor.
As we said in Section~\ref{sect:MPS}, MPS uses a cellular system that enables the language developer to arrange the children and properties of an AST node in a table-like manner.
Cells of different types are supported by MPS --- for storing property values, references to child nodes, and keywords (string literals), and also cells that influence layout (e.g., indentation).
The specific goal of this phase is to create the Editor aspect for each language concept (element), such that all attributes of the concept --- name, properties, children --- are projected using the respective cell types.

The main problem that we had to address is the absence of information about the code layout and whitespace in the ANTLR grammar of an input language.
Rules forming the grammar only tell what the syntax tree looks like and how the program code is decomposed into AST nodes.
Here we present a solution that is only partially automated for reasons explained below.

We experimented with several heuristics to derive a useful code layout, but all of them produced rather suboptimal results.
However, we observed that the most tedious and error-prone step in the manual definition of the Editor aspect is the creation of cells for all literals (keywords), properties, children, and other fields of a given concept that should appear in its visual representation.
This step can be very easily automated.

Our solution that we implemented in the current version of the \textsc{Ingrid} method is to create all the cells and place them in a single row.
The resulting basic layout is illustrated on the example of the \mpsconcept{Element{\_}1} concept in the upper left corner of Figure~\ref{fig:EDITORADJUST}.
Further adjustments of the layout, such as indentation and line breaks, can be done very efficiently in the MPS IDE.
The bottom right part of Figure~\ref{fig:EDITORADJUST} shows a fully customized layout for the \mpsconcept{Element{\_}1} concept.

\begin{figure*}[ht]
	\centering
	\includegraphics[scale=0.55]{./images/editor_adjustment.png}
	\caption{Editor aspect of the \mpsconcept{Element{\_}1} concept}
	\label{fig:EDITORADJUST}
\end{figure*}

Based on our experience, it takes a very short time to manually adjust the layout into a form much better than any fully automated heuristic could achieve.
We expect that a typical user of the \textsc{Ingrid} method will be able to use the projectional editor in MPS quite efficiently.
More details are provided in Section~\ref{sect:EVAL}, including our experience with several mainstream programming languages.

We have found that the combination of two steps, (1) automated placing of cells into a single row and (2) subsequent manual adjustment of the layout, is a very fast and efficient way of creating a nice code layout.
Nevertheless, we plan to work on a more automated approach to the definition of Editor aspects in the future, probably using some techniques of machine learning.
The key idea would be to derive the code layout from a set of valid input source files.
We believe that the user would have to adjust the editor manually a little bit even in this case, because the results of a machine learning-based approach would not be perfect.

\subsection{Phase 4: Text Generation}
\label{sect:TEXTGENDEF}

The purpose of the last phase of the MPS language construction is to generate the TextGen aspect for each concept.
We use the \mpsconcept{Element{\_}1} concept again for illustration.
Figure~\ref{fig:TEXTGENBASIC} shows a very basic example of BaseLanguage code that can be used as its TextGen aspect.
The code appends all literals, properties, and children of \mpsconcept{Element{\_}1} to the output buffer, and does that in the same order as they appear in the corresponding grammar rule.

\begin{figure}[ht]
\begin{alltt}
\small
\mpstgkeyword{text gen component for concept} \mpstgtarget{Element{\_}1} \{
  \mpstgparam{(context, buffer, node)->void} \{
    \mpstgaction{append} \{\mpstgliteral{<}\};
    \mpstgaction{append} \$\{\mpstgparam{node}.\mpstgnodeprop{\textit{Name}}{\_}\textit{1}\};
    \mpstgaction{append} \{\ \};
    \mpstgaction{append} \$list\{\mpstgparam{node}.\mpstgnodeprop{\textit{Attribute}}{\_}\textit{1}\};
    \mpstgaction{append} \{\mpstgliteral{>}\};
    \mpstgaction{append} \$list\{\mpstgparam{node}.\mpstgnodeprop{\textit{Content}}{\_}\textit{2}\};
    \mpstgaction{append} \{\mpstgliteral{</}\};
    \mpstgaction{append} \$\{\mpstgparam{node}.\mpstgnodeprop{\textit{Name}}{\_}\textit{2}\};
    \mpstgaction{append} \{\mpstgliteral{>}\};
  \}
\}
\end{alltt}
\caption{Basic TextGen aspect for the \mpsconcept{Element{\_}1} concept}
\label{fig:TEXTGENBASIC}
\end{figure}

Like in the case of a projectional editor, the main challenge associated with TextGen is to produce valid code with a reasonable layout.
An implementation of the TextGen aspect must determine properly where to put line breaks, indentation, and other whitespace characters.
For example, in the case of the concept representing an XML element, there must be a space between the element's name and the first attribute, while it is not required between the opening bracket \antlrliteral{\textless} and the name (Figure~\ref{fig:TEXTGENBASIC}).

Our \textsc{Ingrid} method targets mostly text-based languages, for the concepts of which the visual representation in a projectional editor must be almost equivalent to their plain-text representation.
An obvious choice would therefore be to use the same approach for Editor and TextGen.
Nevertheless, since the Editor aspect has to be adjusted manually, we decided to use a different approach for TextGen.
We designed a procedure based on a simple fully automated heuristic that provides surprisingly good results.

The procedure creates the TextGen aspect for a given language concept (AST node) in two steps.
First, it generates a basic variant that inserts spaces in between every two tokens of the textual representation of the concept.
In the second step, which is the core of our heuristic, spaces are eliminated from places where they are not really needed.
The main criterion is whether the generated plain-text output can be safely processed by a parser of the original language (using the ANTLR grammar).
We discuss all the cases here:
\begin{itemize}
	\item When there is a non-alphabetical literal that is used as a token in the grammar, and that might get recognized by the parser without the need for whitespace separators around it, then we can omit the spaces. An example of such literal is '\textless' in SimpleXML.
	\item In the case of an arbitrary string, the whitespace may be omitted when the adjacent literal ends, respectively begins, with a non-alphabetical character.
		This applies especially to the values of properties defined in Structure aspects, such as the name of an XML element.
		Based on this heuristic, redundant spaces inside of quotes will be eliminated, as well as spaces next to semicolons and around brackets.
		On the other hand, it will distinguish language keywords (e.g., \code{function}, \code{var}, \code{in}) from other literals by preserving the space character in between them.
	\item Spaces can be safely omitted also when specific optional child nodes are not present, and in the case of empty lists of child nodes, so that whitespace does not accumulate.
\end{itemize}
A space must be always inserted when two child nodes are next to each other.
Elements from a sequence of child nodes are separated with spaces or line breaks.

\begin{figure}[ht]
\begin{alltt}
\small
\mpstgkeyword{text gen component for concept} \mpstgtarget{Element{\_}1} \{
  \mpstgparam{(context, buffer, node)->void} \{
    \mpstgaction{append} \{\mpstgliteral{<}\};
    \mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{\textit{Name}}{\_}\textit{1}.\textbf{isNotEmpty}) \{
      \mpstgaction{append} \$\{\mpstgparam{node}.\mpstgnodeprop{\textit{Name}}{\_}\textit{1}\};
    \}
    \mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{\textit{Attribute}}{\_}\textit{1}.\textbf{size} > 0) \{
      \mpstgaction{append} \{\ \};
      \mpstgaction{append} \$list\{\mpstgparam{node}.\mpstgnodeprop{\textit{Attribute}}{\_}\textit{1} with  \};
    \}
    \mpstgaction{append} \{\mpstgliteral{>}\};
    \mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{\textit{Content}}{\_}\textit{2}.\textbf{size} > 0) \{
      \mpstgaction{append} \$list\{\mpstgparam{node}.\mpstgnodeprop{\textit{Content}}{\_}\textit{2} with  \};
    \}
    \mpstgaction{append} \{\mpstgliteral{</}\};
    \mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{\textit{Name}}{\_}\textit{2}.\textbf{isNotEmpty}) \{
      \mpstgaction{append} \$\{\mpstgparam{node}.\mpstgnodeprop{\textit{Name}}{\_}\textit{2}\};
    \}
    \mpstgaction{append} \{\mpstgliteral{>}\};
  \}
\}
\end{alltt}
\caption{Full TextGen aspect for the \mpsconcept{Element{\_}1} concept}
\label{fig:TEXTGENFINAL}
\end{figure}

\begin{figure}[ht]
\begin{alltt}
\small
\mpstgkeyword{if} (\mpstgparam{node}.\mpstgnodeprop{\textit{Content}}{\_}\textit{2}.\textbf{size} > 0) \{
  \mpstgaction{append} \textcolor{Blue}{\textbf{\textbackslash{}n;}}
  \mpstgaction{indent buffer;}
  \mpstgkeyword{with indent} \{
    \mpstgaction{append} \$list\{\mpstgparam{node}.\mpstgnodeprop{\textit{Content}}{\_}\textit{2} with  \};
  \}
  \mpstgaction{append} \textcolor{Blue}{\textbf{\textbackslash{}n;}}
\} 
\end{alltt}
\caption{Fragment of the TextGen aspect of \mpsconcept{Element{\_}1} with adjusted indentation}
\label{fig:TEXTGENADJUSTED}
\end{figure}

Figure~\ref{fig:TEXTGENFINAL} contains the complete, automatically generated, TextGen aspect for the \mpsconcept{Element{\_}1} concept, which represents the full SimpleXML element.
Such code may be adjusted manually very easily in order to produce nicely indented XML documents.
We only need to wrap the \mpsconcept{Content{\_}2} child node with indentation and change the sequence separator to a new line character.
A fragment of the resulting adjusted aspect (code) is in Figure~\ref{fig:TEXTGENADJUSTED}.

\subsection{Remarks about Grammars}

During our work on the \textsc{Ingrid} method, we have also observed that practical usability of a resulting MPS language depends on the specific manner in which the input ANTLR grammar is defined.
We discuss several issues and our workarounds in this section.

\paragraph{Adjusting grammars.}
In some cases, a small adjustment of the input grammar before the run of \textsc{Ingrid} might yield a better and more useful MPS language.
Here we illustrate this on two simple examples.

For the first example, we use the definition of an XML attribute in Figure~\ref{fig:XMLATTRIB}, which is taken from the original XML grammar.
The lexer rule \antlrlexerrule{STRING} actually says that quotes make a part of the attribute's value.
If the MPS language would faithfully match the grammar, the user would have to always input the leading and trailing quote together with the value in the projectional editor.

\begin{figure}[ht]
\begin{framed}
\begin{alltt}
  \antlrparserrule{attrib} : \antlrlexerrule{Name} \antlrliteral{=} \antlrlexerrule{STRING} ;
  \antlrlexerrule{STRING} : \antlrliteral{"} \antlrregex{~["]*} \antlrliteral{"}
         | \antlrliteral{\textbackslash'} \antlrregex{~[']*} \antlrliteral{\textbackslash'} ;
\end{alltt}
\end{framed}
\caption{Definition of an XML attribute}
\label{fig:XMLATTRIB}
\end{figure}

In our SimpleXML language, we adjusted the grammar easily in a way that we show in Figure~\ref{fig:XMLADJUST}.
We turned quotes into literals, ensuring that they will only appear in the projectional editor as string constant cells.

\begin{figure}[ht]
\begin{framed}
\begin{alltt}
  \antlrparserrule{attrib} : \antlrlexerrule{Name} \antlrliteral{="} \antlrlexerrule{TEXT1} \antlrliteral{"}
         | \antlrlexerrule{Name} \antlrliteral{=\textbackslash'} \antlrlexerrule{TEXT2} \antlrliteral{\textbackslash'} ;
  \antlrlexerrule{TEXT1} : \antlrregex{~["]*} ;
  \antlrlexerrule{TEXT2} : \antlrregex{~[']*} ;
\end{alltt}
\end{framed}
\caption{Adjusted fragment of the SimpleXML grammar}
\label{fig:XMLADJUST}
\end{figure}

The second example relates to a fragment of the JavaScript language, also known as ECMAScript.
%\footnote{https://github.com/antlr/grammars-v4/blob/master/ecmascript/ECMAScript.g4}.
Every statement in JavaScript needs to be terminated by a semicolon, newline, end of the file, or end of the block --- see the Figure~\ref{fig:JAVASCRIPTSTMT}, which contains the corresponding fragment of the ECMAScript grammar.
Since there are multiple options, a user would have to select one of them for each statement in the editor.
Technically, every language concept representing a statement would contain one child node of the interface type \mpsinterface{IEos}, which has to be assigned a proper object (i.e., an AST node corresponding to one of the options listed above).

\begin{figure}[ht]
\begin{framed}
\begin{alltt}
\small
\antlrparserrule{eos} : \antlrlexerrule{SemiColon} | \antlrlexerrule{EOF} | {lineBreakAhead()}? 
    | {{\_}input.LT(1).type() == \antlrlexerrule{CloseBrace}}? ;
\textcolor{gray}{// example reference to the eos rule}
\antlrparserrule{breakStmt} : \antlrlexerrule{Break} \antlrlexerrule{Identifier}? \antlrparserrule{eos} ;
\end{alltt}
\end{framed}
\caption{Statements in JavaScript}
\label{fig:JAVASCRIPTSTMT}
\end{figure}

Since MPS can differentiate between statements on the AST level, there is no need for an explicit separator.
For example, we can put each statement on a distinct line, as is usual for JavaScript code, and keep just the semicolon as a fixed literal in the projectional editor.
The \antlrparserrule{eos} rule must be changed to this form: \antlrparserrule{eos} : \antlrliteral{;}.
A small adjustment of the grammar, like this one, is a very quick solution and definitely makes the generated language more usable in MPS.
On the other hand, this adjustment of the original ANTLR grammar, performed just for the purpose of creating the MPS language, changes the grammar in such a way that it does not precisely describe the standard JavaScript language.
To ensure compatibility with JavaScript, the TextGen aspect in the MPS language should put the semicolon after each statement in the plain-text representation of a program code.

\paragraph{Breaking original grammars and parsers.}
We also want to point out a general problem with grammar adjustments that we encountered, and which all potential users of the \textsc{Ingrid} method should be aware of.
The original ANTLR grammar for an input language can be very easily changed in a way that, at first, seems harmless and valid inside MPS, but the parser generated out of the adjusted grammar stops accepting the original language.
In particular, adjustments needed to improve the resulting MPS language can break down the parser.

This problem has two main causes: (1) low-level implementation details of token matching in the ANTLR parser and (2) usage of ANTLR grammars for a different purpose (that was not expected) in our project.
Since, for practical reasons, it is very important that a parser works even for documents (programs) written according to the original grammar, adjustments have to be performed very carefully by the users of \textsc{Ingrid}.